The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-07-15 script/en/chapters/aboutcryptool.tex
CT script: Spellcheck in order to have consistent American English

Because of didactical reasons the \hyperlink{Kapitel_PaperandPencil}
{second chapter} gives an exhaustive overview
about {\bf paper and pencil encryption methods}.

@@

Whereas the \textit{eLearning program} CrypTool\index{CrypTool} motivates and
teaches you how to use cryptography in practice, the \textit{script} provides
those interested in the subject with a deeper understanding of the mathematical
algorithms used -- trying to do it in an instructive way.
If you are already a little bit familiar with this field of knowledge you can
gain a fast overview about the functions delivered by CrypTool looking at the
{\bf menu tree} (see appendix \ref{s:appendix-menutree}).

@@

\enlargethispage{12pt}
As with the eLearning program CrypTool\index{CrypTool}, the quality of the 
script is enhanced by your suggestions and ideas for improvement. 
We look forward to your feedback.

>>>>

Because of didactic reasons the \hyperlink{Kapitel_PaperandPencil}
{second chapter} gives an exhaustive overview
about {\bf paper and pencil encryption methods}.

@@

Whereas the \textit{e-learning program} CrypTool\index{CrypTool} motivates and
teaches you how to use cryptography in practice, the \textit{script} provides
those interested in the subject with a deeper understanding of the mathematical
algorithms used -- trying to do it in an instructive way.
If you are already a little bit familiar with this field of knowledge you can
gain a fast overview about the functions delivered by CrypTool looking at the
{\bf menu tree} (see appendix \ref{s:appendix-menutree}).

@@

\enlargethispage{12pt}
As with the e-learning program CrypTool\index{CrypTool}, the quality of the 
script is enhanced by your suggestions and ideas for improvement. 
We look forward to your feedback.

### 2009-07-15 script/en/chapters/authors.tex
CT script: Spellcheck in order to have consistent American English

\item[\textrm{[Natali1997]}] \index{Natali 1997}
    Directed by Vincenzo Natali, \\
    {\em Cube}, Mehra Meh Film, 1997. \\
    In this Canadian low-budget-movie 7 complete strangers of widely varying
    personality characteristics are involuntarily placed in an kafkaesque maze
    of cubical rooms containing deadly traps.\\
    To get out the persons have to move through these rooms. To find out
    which rooms are dangerous, mathematics is crucial: Each cubic room has
    at its entrance a numerical marking consisting of three sets of three
    digits. Frist they deduce that all rooms marked at their entrance with
    at least one prime number are trapped. Later it comes out that a trapped
    room can also be marked by a number which is a power of a prime (so 
    traps are $p^n$, e.g. $128=2^7$ or $101 = 101^1 = prime$, but not
    $517 = 11*47$).\\

@@

\item[\textrm{[Vidal2006]}] \index{Vidal 2006}
    Agustin Sanchez Vidal, \\
    {\em Kryptum}, Dtv, 2006. \\
    The first novel of the Spanish professor of art history has some
    similarities with Dan Brown's ``The Da Vinci Code'' from 2003, but allegedly
    Vidal started his writing of the novel already in 1996. Vidal's novel is
    a mixture between historic adventure and mystery thriller.
    It was a huge success in Spain and Germany.
    There is currently no English version available.\\
    In the year 1582 Raimundo Randa is waiting to be condemned to death -- he
    was all life long trying to solve a mystery.
    This mystery is about a pergament with cryptic characters, where a unique
    power is behind.
    Around 400 years later the American scientist Sara Toledano is fascinated
    by this power until she vanishes in Antigua.
    Her colleague, the cryptographer David Calderon, and her daughter Rachel
    are searching for her and simultaneously they try to solve the code.
    But also secret organizations like the NSA chase after the secret
    of the ``last key''. They don't hesitate to kill for it.\\

@@

Kid books with collections of simpler cryptographic encryption methods, 
prepared in a didactical and exciting manner are in the following list
(please send us similar English kid books, because at the moment
our list contains only German kid books):

@@

\item[\textrm{[Flessner2004]}] \index{Flessner 2004}
    Bernd Flessner, \\
    {\em The Three Investigators: Manual for Secret Messages (original
    title: Die 3 ???: Handbuch Geheimbotschaften)},
    Kosmos, 2004. \\
    On 127 pages you learn in an easy and exciting manner, structured by
    the method types, which secret languages (like the one of the Navajo
    indians or dialects) and which secret writings (real encryption or
    hiding via technical or linguistic steganography) existed and how simple
    methods can be decrypted.\\
    The author tells where in history the methods were used and in which
    novel authors used encryption methods [like in Edgar Allan Poe's
    ``The Gold Bug'', like with Jules Verne's hero Mathias Sandorf or like
    with Astrid Lindgren's master detective Blomquist who used the ROR language
    (similar inserting ciphers are the spoon or the B language)].\\
    This is a didactically excellent introduction for younger teens.\\

@@

If you know of further literature and movies, where cryptography has a major
role or if you know of further books, which address cryptography in a
didactical and for children adequate way, then we would be very glad if
you could send us the exact book title and a short explanation
about the book's content. Thanks a lot.

@@

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-2_Fermat-factorization-How-far-1}
\caption{Fermat factorisation using the third binomial formula\vspace{1ex}} 
\label{NT_Fig_C5.2_Fermat-factorization-How-far-1}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-2_Fermat-factorization-How-far-2}
\caption{Fermat factorisation: Detecting squares\vspace{1ex}} 
\label{NT_Fig_C5.2_Fermat-factorization-How-far-2}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-3_PollardRho}
\caption{Pollard's rho factorisation: Floyd's cycle finding algorithm\vspace{1ex}} 
\label{NT_Fig_C5.3_PollardRho}
\end{center}
\end{figure}

>>>>

\item[\textrm{[Natali1997]}] \index{Natali 1997}
    Directed by Vincenzo Natali, \\
    {\em Cube}, Mehra Meh Film, 1997. \\
    In this Canadian low-budget-movie 7 complete strangers of widely varying
    personality characteristics are involuntarily placed in an kafkaesque maze
    of cubical rooms containing deadly traps.\\
    To get out the persons have to move through these rooms. To find out
    which rooms are dangerous, mathematics is crucial: Each cubic room has
    at its entrance a numerical marking consisting of three sets of three
    digits. First they deduce that all rooms marked at their entrance with
    at least one prime number are trapped. Later it comes out that a trapped
    room can also be marked by a number which is a power of a prime (so 
    traps are $p^n$, e.g. $128=2^7$ or $101 = 101^1 = prime$, but not
    $517 = 11*47$).\\

@@

\item[\textrm{[Vidal2006]}] \index{Vidal 2006}
    Agustin Sanchez Vidal, \\
    {\em Kryptum}, Dtv, 2006. \\
    The first novel of the Spanish professor of art history has some
    similarities with Dan Brown's ``The Da Vinci Code'' from 2003, but allegedly
    Vidal started his writing of the novel already in 1996. Vidal's novel is
    a mixture between historic adventure and mystery thriller.
    It was a huge success in Spain and Germany.
    There is currently no English version available.\\
    In the year 1582 Raimundo Randa is waiting to be condemned to death -- he
    was all life long trying to solve a mystery.
    This mystery is about a parchament with cryptic characters, where a unique
    power is behind.
    Around 400 years later the American scientist Sara Toledano is fascinated
    by this power until she vanishes in Antigua.
    Her colleague, the cryptographer David Calderon, and her daughter Rachel
    are searching for her and simultaneously they try to solve the code.
    But also secret organizations like the NSA chase after the secret
    of the ``last key''. They don't hesitate to kill for it.\\

@@

Kid books with collections of simpler cryptographic encryption methods, 
prepared in a didactic and exciting manner are in the following list
(please send us similar English kid books, because at the moment
our list contains only German kid books):

@@

\item[\textrm{[Flessner2004]}] \index{Flessner 2004}
    Bernd Flessner, \\
    {\em The Three Investigators: Manual for Secret Messages (original
    title: Die 3 ???: Handbuch Geheimbotschaften)},
    Kosmos, 2004. \\
    On 127 pages you learn in an easy and exciting manner, structured by
    the method types, which secret languages (like the one of the Navajo
    Indians or dialects) and which secret writings (real encryption or
    hiding via technical or linguistic steganography) existed and how simple
    methods can be decrypted.\\
    The author tells where in history the methods were used and in which
    novel authors used encryption methods [like in Edgar Allan Poe's
    ``The Gold Bug'', like with Jules Verne's hero Mathias Sandorf or like
    with Astrid Lindgren's master detective Blomquist who used the ROR language
    (similar inserting ciphers are the spoon or the B language)].\\
    This is a didactically excellent introduction for younger teens.\\

@@

If you know of further literature and movies, where cryptography has a major
role or if you know of further books, which address cryptography in a
didactic and for children adequate way, then we would be very glad if
you could send us the exact book title and a short explanation
about the book's content. Thanks a lot.

@@

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-2_Fermat-factorization-How-far-1}
\caption{Fermat factorization using the third binomial formula\vspace{1ex}} 
\label{NT_Fig_C5.2_Fermat-factorization-How-far-1}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-2_Fermat-factorization-How-far-2}
\caption{Fermat factorization: Detecting squares\vspace{1ex}} 
\label{NT_Fig_C5.2_Fermat-factorization-How-far-2}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-3_PollardRho}
\caption{Pollard's rho factorization: Floyd's cycle finding algorithm\vspace{1ex}} 
\label{NT_Fig_C5.3_PollardRho}
\end{center}
\end{figure}

### 2009-07-15 script/en/chapters/cryptomethods.tex
CT script: Spellcheck in order to have consistent American English

The purpose of encryption \index{Encryption} is to change data in such a way
that only an authorised recipient is able to reconstruct the plaintext. This
allows us to transmit data without worrying about it getting into unauthorised
hands. Authorised recipients possess a piece of secret information -- called the
key -- which allows them to decrypt the data while it remains hidden from
everyone else.\par \vskip + 3pt

@@

% --------------------------------------------------------------------------
\section[Symmetric encryption]
{Symmetric encryption\footnotemark}
  \footnotetext{%
    With CrypTool\index{CrypTool} you can execute the following modern
    symmetric encryption algorithms 
    (using the menu path {\bf Crypt \textbackslash{} Symmetric (modern)}): \\
    IDEA, RC2, RC4, DES (ECB), DES (CBC), Triple-DES (ECB), Triple-DES (CBC),
    MARS (AES candidate), RC6 (AES candidate), Serpent (AES candidate), 
    Twofish (AES candidate), Rijndael (official AES algorithm).
  }

For {\em symmetric} encryption \index{Encryption!symmetric} sender and
recipient must be in possession of a common (secret) key which they have
exchanged before actually starting to communicate. The sender uses this
key to encrypt the message and the recipient uses it to decrypt it.
\par \vskip + 3pt

@@

Similar cipher challenges are there for asymmetric algorithms (please see chapter \ref{NoteFactorisation}).

@@

In this case the data is encrypted using symmetric procedures: the key is a
session key\index{Session key} generated by the sender
randomly\footnote{%
An important part of cryptographically secure techniques is to generate 
random numbers. Within CrypTool\index{CrypTool} you can check out
different random number generators using the menu path
{\bf Indiv. Procedures \textbackslash{} Generate Random Numbers}. 
Using the menu path {\bf Analysis \textbackslash{} Analyse Randomness}
you can apply different test methods for random data to binary documents. \\
Up to now CrypTool has concentrated on cryptographically strong 
pseudo random number generators. Only the integrated Secude\index{SECUDE IT Security}
generator involves a "pure" random source. 
}\index{Random}
that is only used for this message.
This session key is then encrypted using the asymmetric procedure and
transmitted to the recipient together with the message. Recipients can determine
the session key using their secret keys and then use the session key to encrypt
the message. In this way, we can benefit from the easy key management
\index{Key management} of asymmetric procedures and encrypt large quantities of data
quickly and efficiently using symmetric procedures.

>>>>

The purpose of encryption \index{Encryption} is to change data in such a way
that only an authorized recipient is able to reconstruct the plaintext. This
allows us to transmit data without worrying about it getting into unauthorized
hands. Authorized recipients possess a piece of secret information -- called the
key -- which allows them to decrypt the data while it remains hidden from
everyone else.\par \vskip + 3pt

@@

% --------------------------------------------------------------------------
\section[Symmetric encryption]
{Symmetric encryption\footnotemark}
  \footnotetext{%
    With CrypTool\index{CrypTool} you can execute the following modern
    symmetric encryption algorithms 
    (using the menu path {\bf Crypt \textbackslash{} Symmetric (modern)}): \\
    IDEA, RC2, RC4, DES (ECB), DES (CBC), Triple-DES (ECB), Triple-DES (CBC),
    MARS (AES candidate), RC6 (AES candidate), Serpent (AES candidate), 
    Twofish (AES candidate), Rijndael (official AES algorithm).
  }
\nopagebreak
For {\em symmetric} encryption \index{Encryption!symmetric} sender and
recipient must be in possession of a common (secret) key which they have
exchanged before actually starting to communicate. The sender uses this
key to encrypt the message and the recipient uses it to decrypt it.
\par \vskip + 3pt

@@

Similar cipher challenges are there for asymmetric algorithms (please see chapter \ref{NoteFactorization}).

@@

In this case the data is encrypted using symmetric procedures: the key is a
session key\index{Session key} generated by the sender
randomly\footnote{%
An important part of cryptographically secure techniques is to generate 
random numbers. Within CrypTool\index{CrypTool} you can check out
different random number generators using the menu path
{\bf Indiv. Procedures \textbackslash{} Generate Random Numbers}. 
Using the menu path {\bf Analysis \textbackslash{} Analyze Randomness}
you can apply different test methods for random data to binary documents. \\
Up to now CrypTool has concentrated on cryptographically strong 
pseudo random number generators. Only the integrated Secude\index{SECUDE IT Security}
generator involves a "pure" random source. 
}\index{Random}
that is only used for this message.
This session key is then encrypted using the asymmetric procedure and
transmitted to the recipient together with the message. Recipients can determine
the session key using their secret keys and then use the session key to encrypt
the message. In this way, we can benefit from the easy key management
\index{Key management} of asymmetric procedures and encrypt large quantities of data
quickly and efficiently using symmetric procedures.

### 2009-07-15 script/en/chapters/digitalsignatures.tex
CT script: Spellcheck in order to have consistent American English

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. Because a signature is about
as long as the straight datastream to be signed, the procedure we just described
has in practice, however, a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we apply
a cryptographic hash function\footnote{%
Hash functions\index{Hash function} are implemented within 
CrypTool\index{CrypTool} at several places.\\
Using menus {\bf Individual Procedures \textbackslash{} Hash} and
            {\bf Analysis \textbackslash{} Hash}
you have the possibilities
% hier die items nicht einruecken!
\begin{list}{\textbullet}{\leftmargin10pt\addtolength{\itemsep}{-1.0\baselineskip}}
%\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hash value,\\
\item to calculate a key from a password according to the PKCS\#5 
      standard\index{PKCS\#5}, \\
\item to calculate HMACs from a text and a secret key, and\\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions\index{Collision}.
\end{list}
} to the document -- before signing. The output of the hash function will 
then be signed.

>>>>

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. Because a signature is about
as long as the straight data stream to be signed, the procedure we just described
has in practice, however, a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we apply
a cryptographic hash function\footnote{%
Hash functions\index{Hash function} are implemented within 
CrypTool\index{CrypTool} at several places.\\
Using menus {\bf Individual Procedures \textbackslash{} Hash} and
            {\bf Analysis \textbackslash{} Hash}
you have the possibilities
% hier die items nicht einruecken!
\begin{list}{\textbullet}{\leftmargin10pt\addtolength{\itemsep}{-1.0\baselineskip}}
%\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hash value,\\
\item to calculate a key from a password according to the PKCS\#5 
      standard\index{PKCS\#5}, \\
\item to calculate HMACs from a text and a secret key, and\\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions\index{Collision}.
\end{list}
} to the document -- before signing. The output of the hash function will 
then be signed.

### 2009-07-15 script/en/chapters/ellipticcurves.tex
CT script: Spellcheck in order to have consistent American English

Nevertheless, thin clients like smart cards usually have to store the (long)
secret key and have to process a digital signature rather than verify one.
Therefore, there is a clear advantage in using ECC in terms of efficiency.
\par
\smallskip
Nowadays, the major problem with ECC-implementations is the lack of standardization.
There is only one way to implement RSA, but there are many ways for ECC: One can work with
different sets of numbers, different (elliptic) curves --- described by parameters\footnote{%
see chapter \ref{ECC-Crypto}
} --- ,
and a variety of representations of the elements on the curve. Each choice has its
advantages and disadvantages, and one can certainly construct the most efficient for
each application. However, this causes problems in interoperability. But if all
ECC-tools should be able to communicate with each other, they will have to support
all different algorithms, which might put the advantage of efficient computation and
the need of less storage capacity to the contrary.

Therefore, international standardization organizations like IEEE (P1363),
ASC (ANSI X9.62, X9.63), ISO/IEC as well as major players like RSA labs or
Certicom have recently started standardization initiatives. While the IEEE
only describes the different implementations, the ASC has explicitly stated
10 elliptic curves and recommends their usage. The advantage of the ASC
approach is that one needs only a single byte to indicate which curve is
meant. However, it is not yet clear whether the ASC-curves will become a de
facto standard.

@@

Anyway that's enough about the role of elliptic curves in pure mathematics. 
In 1985 Neal Koblitz\index{Koblitz, Neal} and Victor Miller\index{Miller, Victor}
independently suggested using elliptic curves in cryptography. Elliptic 
curves have thus also found a concrete practical application. 
Another interesting area of application for elliptic curves is for
factorising whole numbers \index{Factorisation} (the RSA cryptographic system
is based on the \index{Complexity} difficulty/complexity of finding prime
factors of an extremely large number;  compare section \ref{SecurityRSA}.).
In this area, procedures based on elliptic curves have been investigated and
used since 1987 (compare section \ref{ECC-Factorisation}). \\
There are also prime number tests\index{Prime number!test} based on elliptic
curves.
% (s. Anmerkung U. Kuehn)

Elliptic curves are used differently in the various areas. Encryption
procedures based on elliptic curves are based on the difficulty of a problem
known as elliptic curve discrete logarithm\index{Logarithm problem!discrete}.
The factorisation of whole numbers uses the fact that a large number of
elliptic curves can be generated for a natural composite number $n$ with
several prime factors; however, these curves are not then groups for composite
$n$. More information about this can be found under the chapter \ref{ECC-Factorisation}.

@@

% -----------------------------------------------------------------------------
\hypertarget{faktell}{}
\section{Factorisation using elliptic curves} \label{ECC-Factorisation}

There are factorisation%
\footnote{Especially John M. Pollard \index{Pollard, John M.} was involved
in the development of many different factorisation algorithms; also at
factorisation with ECC he was one of the leading heads. As an employee
of British Telekom he never published much. At the RSA data Security Conference
in 1999 he was awarded for his ``outstanding contributions in mathematics'':
\href{http://www.eff.org/Privacy/Crypto_misc/DESCracker/HTML/19990118_rsa_awards.html}
{\texttt{http://www.eff.org/Privacy/Crypto\_misc/DESCracker/HTML/19990118\_rsa\_awards.html}}.
}
algorithms based on elliptic curves%
\footnote{In 1987 H.W. Lenstra published
a factorization algorithm, based on elliptic curves (see \cite{Lenstra1987}).
The biggest compound number currently 
factorised\index{Factorisation!factoring records} with elliptic curves is 
the number $ 628^{59}-1, $ which has 55 decimal digits. It was
found Oct. 6th, 2001 by M. Izumi 
(See \hyperlink{Lenstra2}{ECMNET}\index{ECMNET}).
}. 
More precisely, these procedures exploit the fact that elliptic curves can
be defined over ${\mathbb Z}_n$ ($n$ composite number). Elliptic curves 
over ${\mathbb Z}_n$ do not form a group, because not every point on such 
an elliptic curve has an inverse point. This is connected with the fact 
that - if $n$ is a composite
number - there exist elements in ${\mathbb Z}_n$ that do not have an inverse
with respect to multiplication mod $n$. In order to add two points on an
elliptic curve over ${\mathbb Z}_n$, we can calculate in the same way as on
elliptic curves over ${\mathbb Z}_p$. Addition of two points (on an elliptic
curve over ${\mathbb Z}_n$), however, fails if and only if a factor of $n$ has
been found. The reason for this is that the procedure for adding points on
elliptic curves gives elements in ${\mathbb Z}_n$ and calculates the inverse
elements for these (with respect to multiplication mod $n$) in ${\mathbb Z}_n$.
The extended \index{Euclidean algorithm} Euclidean algorithm is used here. If
the addition of two points (that lie of an elliptic curve over ${\mathbb Z}_n$)
gives an element in ${\mathbb Z}_n$ that does not have an inverse element in
${\mathbb Z}_n$, then the extended Euclidean algorithm delivers a genuine factor
of $n$.

Factorisation using elliptic curves thus principally works as follows: 
Random curves over ${\mathbb Z}_n$ are selected, as well as random points
(that lie on this curve) and add them; you thus obtain points that also
lie on the curve or find a factor of $n$. 
Factorisation algorithms based on elliptic curves
therefore work probabilistically. The opportunity of defining large number of
elliptic curves over ${\mathbb Z}_n$ allows you to increase the probability of
finding two points which you can add to obtain a factor of $n$. These procedures
are therefore highly suitable for parallelisation.

@@

   \item \hypertarget{Lenstra2}{}
         An informative web page about factorisation with elliptic curves, \\
         \href{http://www.loria.fr/~zimmerma/records/ecmnet.html}
	 {\texttt{http://www.loria.fr/\~{}zimmerma/records/ecmnet.html}} \\
         It contains literature related to the topic factorisation with 
	 elliptic curves as well as links to other web page.

>>>>

Nevertheless, thin clients like smart cards usually have to store the (long)
secret key and have to process a digital signature rather than verify one.
Therefore, there is a clear advantage in using ECC in terms of efficiency.
\par
\smallskip
Nowadays, the major problem with ECC implementations is the lack of standardization.
There is only one way to implement RSA, but there are many ways for ECC: One can work with
different sets of numbers, different (elliptic) curves --- described by parameters\footnote{%
see chapter \ref{ECC-Crypto}
} --- ,
and a variety of representations of the elements on the curve. Each choice has its
advantages and disadvantages, and one can certainly construct the most efficient for
each application. However, this causes problems in interoperability. But if all
ECC-tools should be able to communicate with each other, they will have to support
all different algorithms, which might put the advantage of efficient computation and
the need of less storage capacity to the contrary.

Therefore, international standardization organizations like IEEE (P1363),
ASC (ANSI X9.62, X9.63), ISO/IEC as well as major players like RSA labs or
Certicom have recently started standardization initiatives. While the IEEE
only describes the different implementations, the ASC has explicitly stated
10 elliptic curves and recommends their usage. The advantage of the ASC
approach is that one needs only a single byte to indicate which curve is
meant. However, it is not yet clear whether the ASC curves will become a de
facto standard.

@@

Anyway that's enough about the role of elliptic curves in pure mathematics. 
In 1985 Neal Koblitz\index{Koblitz, Neal} and Victor Miller\index{Miller, Victor}
independently suggested using elliptic curves in cryptography. Elliptic 
curves have thus also found a concrete practical application. 
Another interesting area of application for elliptic curves is for
factorizing whole numbers \index{Factorization} (the RSA cryptographic system
is based on the \index{Complexity} difficulty/complexity of finding prime
factors of an extremely large number;  compare section \ref{SecurityRSA}.).
In this area, procedures based on elliptic curves have been investigated and
used since 1987 (compare section \ref{ECC-Factorization}). \\
There are also prime number tests\index{Prime number!test} based on elliptic
curves.
% (s. Anmerkung U. Kuehn)

Elliptic curves are used differently in the various areas. Encryption
procedures based on elliptic curves are based on the difficulty of a problem
known as elliptic curve discrete logarithm\index{Logarithm problem!discrete}.
The factorization of whole numbers uses the fact that a large number of
elliptic curves can be generated for a natural composite number $n$ with
several prime factors; however, these curves are not then groups for composite
$n$. More information about this can be found under the chapter \ref{ECC-Factorization}.

@@

% -----------------------------------------------------------------------------
\hypertarget{faktell}{}
\section{Factorization using elliptic curves} \label{ECC-Factorization}

There are factorization%
\footnote{Especially John M. Pollard \index{Pollard, John M.} was involved
in the development of many different factorization algorithms; also at
factorization with ECC he was one of the leading heads. As an employee
of British Telekom he never published much. At the RSA data Security Conference
in 1999 he was awarded for his ``outstanding contributions in mathematics'':
\href{http://www.eff.org/Privacy/Crypto_misc/DESCracker/HTML/19990118_rsa_awards.html}
{\texttt{http://www.eff.org/Privacy/Crypto\_misc/DESCracker/HTML/19990118\_rsa\_awards.html}}.
}
algorithms based on elliptic curves%
\footnote{In 1987 H.W. Lenstra published
a factorization algorithm, based on elliptic curves (see \cite{Lenstra1987}).
The biggest compound number currently 
factorized\index{Factorization!factoring records} with elliptic curves is 
the number $ 628^{59}-1, $ which has 55 decimal digits. It was
found Oct. 6th, 2001 by M. Izumi 
(See \hyperlink{Lenstra2}{ECMNET}\index{ECMNET}).
}. 
More precisely, these procedures exploit the fact that elliptic curves can
be defined over ${\mathbb Z}_n$ ($n$ composite number). Elliptic curves 
over ${\mathbb Z}_n$ do not form a group, because not every point on such 
an elliptic curve has an inverse point. This is connected with the fact 
that - if $n$ is a composite
number - there exist elements in ${\mathbb Z}_n$ that do not have an inverse
with respect to multiplication mod $n$. In order to add two points on an
elliptic curve over ${\mathbb Z}_n$, we can calculate in the same way as on
elliptic curves over ${\mathbb Z}_p$. Addition of two points (on an elliptic
curve over ${\mathbb Z}_n$), however, fails if and only if a factor of $n$ has
been found. The reason for this is that the procedure for adding points on
elliptic curves gives elements in ${\mathbb Z}_n$ and calculates the inverse
elements for these (with respect to multiplication mod $n$) in ${\mathbb Z}_n$.
The extended \index{Euclidean algorithm} Euclidean algorithm is used here. If
the addition of two points (that lie of an elliptic curve over ${\mathbb Z}_n$)
gives an element in ${\mathbb Z}_n$ that does not have an inverse element in
${\mathbb Z}_n$, then the extended Euclidean algorithm delivers a genuine factor
of $n$.

Factorization using elliptic curves thus principally works as follows: 
Random curves over ${\mathbb Z}_n$ are selected, as well as random points
(that lie on this curve) and add them; you thus obtain points that also
lie on the curve or find a factor of $n$. 
Factorization algorithms based on elliptic curves
therefore work probabilistically. The opportunity of defining large number of
elliptic curves over ${\mathbb Z}_n$ allows you to increase the probability of
finding two points which you can add to obtain a factor of $n$. These procedures
are therefore highly suitable for parallelization.

@@

   \item \hypertarget{Lenstra2}{}
         An informative web page about factorization with elliptic curves, \\
         \href{http://www.loria.fr/~zimmerma/records/ecmnet.html}
	 {\texttt{http://www.loria.fr/\~{}zimmerma/records/ecmnet.html}} \\
         It contains literature related to the topic factorization with 
	 elliptic curves as well as links to other web page.

### 2009-07-15 script/en/chapters/introduction.tex
CT script: Spellcheck in order to have consistent American English

The authors
% \hyperlink{appendix-authors}{(authors)}
% in \ref{s:appendix-authors}
% \hypertarget{appendix-authors}{}\label{s:appendix-authors}
have attempted to describe cryptography for a broad 
audience -- without being mathematically incorrect. We believe that this
didactical pretension is the best way to promote the awareness for IT
security and the readiness to use standardised modern cryptography.
\par \vskip + 15pt

>>>>

The authors
% \hyperlink{appendix-authors}{(authors)}
% in \ref{s:appendix-authors}
% \hypertarget{appendix-authors}{}\label{s:appendix-authors}
have attempted to describe cryptography for a broad 
audience -- without being mathematically incorrect. We believe that this
didactic pretension is the best way to promote the awareness for IT
security and the readiness to use standardized modern cryptography.
\par \vskip + 15pt
